Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating

### 2017-12-01 SubGenome Analysis.Rmd
Kaeppler expression data
Import and use kaeppler (must replace Walley data at bottom of do.R)

Updated reports

Made the log10_ks_cutoff a variable and not a knitr param.

Fixed a few minor issues in knitr reports.

---
title: 'Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating
  subgenome assignments in maize'
params:
  log10_ks_cutoff: 0
output:
  html_document: default
  pdf_document: default
---

@@

You have selected the log10(ks) cutoff value to be **`r params$log10_ks_cutoff`**.  Again, this should ideally separate two peaks in the middle graph.

```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3, fig.cap = cap}
qplot(
  syntelogs.mutated$ks[syntelogs.mutated$ks < 2.5], 
  geom="histogram", binwidth=.05, na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = 10^params$log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of KS values by \nsyntenic block (KS < 2.5)", 
  x="KS value of syntenic block", 
  y="Count (# of genes)"
)

qplot(
  syntelogs.mutated$median_ks[syntelogs.mutated$median_ks < 2.5], 
  geom="histogram", 
  binwidth=.05, 
  na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = 10^params$log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of median KS \nvalues by syntenic block \n(median KS < 2.5)", 
  x="Median KS value of syntenic block", 
  y="Count (# of genes)"
)

qplot(
  log10(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$ks), 
  geom="histogram", 
  binwidth=.01, 
  na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = params$log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of log10 KS \nvalues", 
  x="log10(KS) value of genes", 
  y="Count (# of genes)"
)
cap <- "Fig 2.1: "
```

@@

```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4, fig.align='center'}
# Expression Diffs
grid.newpage()
g <- draw.pairwise.venn(
  area1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2) %>% nrow(),
  area2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2) %>% nrow(),
  cross.area = intersect(expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) < 2),
                         expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) < 2)) %>%
                          nrow(), 
  category = c("Sub1", "Sub2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
```

@@

Paired t-test suggests that there is no difference in means between FPKM of homeologs in sub1 vs. sub2 (p-val > 0.05, so we accept null hypothesis of equal means):
```{r subgenome_fpkm_ttest, echo=FALSE}
t.test(x=expressedPairs$FPKM_mean1, y=expressedPairs$FPKM_mean2, paired = TRUE)
```

>>>>

---
title: 'Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating
  subgenome assignments in maize'
output:
  html_document: default
  pdf_document: default
---

@@

You have selected the log10(ks) cutoff value to be **`r log10_ks_cutoff`**.  Again, this should ideally separate two peaks in the middle graph.

```{r pickKSCutoff, echo=FALSE, fig.width=3, fig.height=3}
qplot(
  syntelogs.mutated$ks[syntelogs.mutated$ks < 2.5], 
  geom="histogram", binwidth=.05, na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = 10^log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of KS values by \nsyntenic block (KS < 2.5)", 
  x="KS value of syntenic block", 
  y="Count (# of genes)"
)

qplot(
  syntelogs.mutated$median_ks[syntelogs.mutated$median_ks < 2.5], 
  geom="histogram", 
  binwidth=.05, 
  na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = 10^log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of median KS \nvalues by syntenic block \n(median KS < 2.5)", 
  x="Median KS value of syntenic block", 
  y="Count (# of genes)"
)

qplot(
  log10(syntelogs.sorghum.v3.1.maize.v4.and.rejected.clean$ks), 
  geom="histogram", 
  binwidth=.01, 
  na.rm=TRUE
) + geom_vline(
  mapping = NULL, 
  data = NULL, 
  xintercept = log10_ks_cutoff, 
  na.rm = FALSE, 
  show.legend = NA
) + labs(
  title="Frequency of log10 KS \nvalues", 
  x="log10(KS) value of genes", 
  y="Count (# of genes)"
)
cap <- "Fig 2.1: "
```

@@

```{r diffExpression_subgenome, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4, fig.height=4, fig.align='center'}
# Expression Diffs
grid.newpage()
a1 = expressedPairs %>% subset((FPKM_mean1/FPKM_mean2) > 2) %>% nrow()
a2 = expressedPairs %>% subset((FPKM_mean2/FPKM_mean1) > 2) %>% nrow()
a3 = nrow(expressedPairs) - a1 - a2
g <- draw.pairwise.venn(
  area1 = a1+a3,
  area2 = a2+a3,
  cross.area = a3, 
  category = c("Sub1", "Sub2"), 
  lty = rep("blank", 2), 
  fill = c("light blue", "green"), 
  alpha = rep(0.5, 2), 
  cat.pos = c(0, 180), 
  scaled = FALSE,
  euler.d = FALSE, 
  sep.dist = 0.3, 
  rotation.degree = 0,
  ind = FALSE
)
grid.arrange(gTree(children=g), top="Genes with 2-fold higher\nexpression by subgenome")
```

@@

Paired (two-sided) t-test suggests that there is no difference in means between FPKM of homeologs in sub1 vs. sub2 (p-val > 0.05, so we fail to reject null hypothesis of equal means):
```{r subgenome_fpkm_ttest, echo=FALSE}
t.test(x=expressedPairs$FPKM_mean1, y=expressedPairs$FPKM_mean2, paired = TRUE)
```
